---
title: "Getting Started with MassKitUtils"
author: "Songbiao Zhu"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with MassKitUtils}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Getting Started with MassKitUtils

MassKitUtils is a comprehensive utility package for R development and data export. This guide will help you get started with the package.

## Installation

```{r eval=FALSE}
# Install from GitHub
devtools::install_github("SongbiaoZhu/MassKitUtils")

# Load the package
library(MassKitUtils)
```

## Quick Start

### 1. Package Management

Ensure required packages are installed:

```{r}
# Load the package first
library(MassKitUtils)

# Install and load packages
ensure_packages(c("dplyr", "ggplot2"))
load_packages(c("dplyr", "ggplot2"))

# Check package versions
check_package_versions(c("dplyr", "ggplot2"))
```

### 2. Project Setup

Create a new R project with standard structure:

```{r eval=FALSE}
# Initialize a new R project
init_r_project()

# Or create directory structure in current directory
setup_analysis_structure()
```

### 3. Data Export

Export data to Excel with formatting:

```{r}
# Create sample data
data <- data.frame(
  Name = c("Alice", "Bob", "Charlie"),
  Age = c(25, 30, 35),
  Score = c(85, 92, 78)
)

# Export to Excel
export_to_excel(data, "sample_data.xlsx", 
                header_style = "colored", 
                table_style = "striped")
```

### 4. Visualization

Create publication-ready plots:

```{r eval=FALSE}
library(ggplot2)

# Create a plot with publication theme
p <- ggplot(mtcars, aes(x = wt, y = mpg, color = factor(cyl))) +
  geom_point() +
  labs(title = "Publication Quality Plot") +
  theme_publication()

# Save with publication quality
save_publication_plot(p, "my_plot.png", 
                     target_width = 8, target_height = 6)
```

## Key Features

- **Package Management**: Intelligent package installation and loading
- **Project Setup**: Standardized project structure creation
- **Data Export**: Excel export with formatting options
- **Visualization**: Publication-ready ggplot2 themes
- **Development Tools**: Standards and templates generation

## Next Steps

- Read the [function documentation](reference/index.html) for detailed usage
- Check out the [tutorials](articles/tutorials.html) for advanced examples
- Visit the [GitHub repository](https://github.com/SongbiaoZhu/MassKitUtils) for the latest updates 